---
title: ""
author: "Group 33"
date: "November 9, 2024"
---

```{r setup, include=FALSE}
# Load necessary libraries
suppressPackageStartupMessages({
  library(tidyverse)
  library(tidymodels)
  theme_set(theme_bw())
  library(patchwork)
  library(fs)
  library(dplyr)
  library(readr)
})
```

# Group member:

-   Anh Phuong Dieu Nguyen (169105483)

-   Minh Tri Tran (169107769)

-   Thi Thao Quyen Tran (169077196)

# Introduction 

## Context

Unnecessary, problematic, and excessive single-use packaging is one of the most offensive and damaging applications of plastic. Despite calls for a “circular economy” approach to plastic packaging, thousands of brands and corporations insist on wrapping their food, household, and personal care products in this polluting and toxic material.

In this report, we will...

## Reasearch Goal 

By the end of this report, we aim to...

# Data Description

## Plastic Pollution Data

.

## Choice of Variables

From this data, we choose the total burned area caused by a fire as our target. To predict this information, we consider several features attributes, including:

-   State in which the fire started, including Arizona, California, Nevada, Oregon.

-   Objective, which indicates the intended goals associated with managing the fire (extinguish or allow to burn).

-   Cause, by human or nature or neither.

-   Shape area, which is the total two-dimensional area enclosed by the fire perimeter polygon.

-   Shape length, which is the total length of the boundary (or perimeter) of the fire polygon.

# Basic Data Cleaning

```{r clean}
# Load dataset
cal_fire <- read.csv("D:/WLU/1st Year/Fall/DATA100/FINAL PROJECT/California_Fire_Perimeters_(all).csv")

# Clean and reorder the dataset according to recycling codes

cal_fire <- cal_fire |> 
  na.omit(cal_fire) |> 
  mutate(CAUSE = case_when(
    CAUSE == 1 | CAUSE == 17 ~ "Natural",
    CAUSE == 14 | is.na(CAUSE) ~ "Unknown",
    CAUSE != 1 | CAUSE != 14 | CAUSE != 17 ~ "Human")) |> 
  mutate(
    ALARM_DATE=ymd_hms(ALARM_DATE),
    CONT_DATE=ymd_hms(CONT_DATE),
    GIS_ACRES=format(GIS_ACRES, scientific=FALSE),
    Shape__Area=format(Shape__Area, scientific=FALSE),
    Shape__Length=format(Shape__Length, scientific=FALSE),
  ) |> 
  rename(BURNED_ACRES=GIS_ACRES,
         SHAPE_AREA=Shape__Area,
         SHAPE_LENGTH=Shape__Length) |> 
  select(!c(COMMENTS,COMPLEX_NAME,IRWINID,FIRE_NUM,COMPLEX_ID))
cal_fire
```

# Exploratory Plot/Table

## Exploratory Plot  1

Shape area vs. Shape length

```{r exploratory plot 1}
exploratory_plot1 <- cal_fire |> 
  ggplot(aes(x=SHAPE_LENGTH, y=SHAPE_AREA, colour=OBJECTIVE)) +
  geom_point()
exploratory_plot1

```

## Exploratory Plot 2

```{r exploratory plot 2}

# Ensure BURNED_ACRES is numeric
cal_fire <- cal_fire |>
  mutate(BURNED_ACRES = as.numeric(BURNED_ACRES))

exploratory_plot2 <- cal_fire |>
  group_by(YEAR_) |>
  summarise(avg_burned_acres = mean(BURNED_ACRES, na.rm = TRUE)) |>
  ggplot(aes(x = YEAR_, y = avg_burned_acres)) +
  geom_col(fill = "orange", alpha = 0.8) +
  labs(title = "Average Burned Acres by Year",
       x = "Year",
       y = "Average Burned Acres") +
  theme_minimal()


exploratory_plot2
```

# Model Plot 

## Model Plot 1
```{r}
ggplot(cal_fire, aes(x = SHAPE_AREA, y = BURNED_ACRES)) +
  geom_point(alpha = 0.5, color = "blue") +
  labs(title = "Burned Acres vs Shape Area", x = "Shape Area", y = "Burned Acres")


```

## Model Plot 2
```{r}
ggplot(cal_fire, aes(x = YEAR_, y = BURNED_ACRES)) +
  geom_point(alpha = 0.5, color = "BLUE") +
  labs(title = "Burned Acres vs Year", x = "Year", y = "Burned Acres")

```

# Exploratory Linear Model

## Exploratory Linear Model 1
```{r}
exploratory_model_1 <- lm(BURNED_ACRES ~ SHAPE_AREA, data = cal_fire)
cat("Exploratory Linear Model 1 Summary:\n")
print(summary(exploratory_model_1))
```

## Exploratory Linear Model 2
```{r}
exploratory_model_2 <- lm(BURNED_ACRES ~ YEAR_, data = cal_fire)
cat("\nExploratory Linear Model 2 Summary:\n")
print(summary(exploratory_model_2))

```

# Final Linear Models
```{r}
final_model <- lm(BURNED_ACRES ~ SHAPE_AREA + YEAR_, data = cal_fire)
cat("\nFinal Linear Model Summary:\n")
print(summary(final_model))
```

# Advanced Model
```{r}
set.seed(123)
train_index <- createDataPartition(cal_fire$BURNED_ACRES, p = 0.7, list = FALSE)
train_data <- cal_fire[train_index, ]
test_data <- cal_fire[-train_index, ]

rf_model <- randomForest(BURNED_ACRES ~ SHAPE_AREA + YEAR_, data = train_data, ntree = 100)


rf_predictions <- predict(rf_model, test_data)


mse <- mean((rf_predictions - test_data$BURNED_ACRES)^2)
r_squared <- 1 - sum((rf_predictions - test_data$BURNED_ACRES)^2) / sum((test_data$BURNED_ACRES - mean(test_data$BURNED_ACRES))^2)

```

# Conclusions

# Limitation

# References

## Model Building



## Predictions and Augment
